---
title: "Assignment_2_data_cleaning"
author: "Vaibhav Bhatt"
date: "15/11/2021"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=TRUE}
# Describing data
```
DATA SUMMARY: 
This data uses "Centers for Disease Control and Prevention’s Behavioral Risk Factor Surveillance System (BRFSS)" to present the information on predicted population in a metropolitan division that is at high risk from COVID-19 and the number of ICU beds in that division.

DATA FIELDS:
 1. MMSA: The name of the metropolitan area, micropolitan area or metropolitan division available in the CDC’s BRFSS
 2. total_percent_at_risk:	The percent of individuals in that area that are at high risk of becoming seriously ill from COVID-19, per CDC’s BRFSS
 3. high_risk_per_icu_bed:	The number of high risk individuals per ICU bed in that area
 4. high_risk_per_hospital: The number of high risk individuals per hospital in that area
 5. icu_beds: The number of ICU beds in the area, based on the Kaiser Family Foundation’s data
 6. hospitals:	The number of hospitals in the area, based on the Kaiser Family Foundation’s data
 7. total_at_risk:	The total number of high risk individuals in the area, per CDC’s BRFSS

DATA FORMAT: Data is downloaded from the data source in .csv format, where values are separated by delimiter "comma". 

```{r, echo=TRUE}
library(readr)
library(tidyverse)
library(dplyr)

# creating a variable that has csv file along with the local file path
local_file <- "/Users/vaibhavbhatt/Downloads/mmsa-icu-beds.csv"

# reading csv data using "readr" package and storing it in a variable "csv_data" using "read_csv"
csv_data <- read_csv(local_file)

# preview of the data loaded from local file into R
head(csv_data)
```

```{r, echo=TRUE}
# Performing two data cleaning functions:
#   1. Find the rows with missing values
#   2. Remove the rows with the missing values from the data

# Check if the data frame or matrix has any missing value
is.na(csv_data)

# removing missing values from csv_data
csv_data <- csv_data[complete.cases(csv_data), ]
# partial_records <- sampledata[complete.cases(sampledata)]
```

```{r, echo=TRUE}
# renaming the column names using dplyr
library(dplyr)

# manipulating column name in the csv_data and storing it in the original variable
csv_data <- csv_data %>%
  rename(patient_per_ICU_bed = high_risk_per_ICU_bed,
         percent_at_risk = total_percent_at_risk,
         )

# Checking if the changes in the column names reflected in the "csv_data"
head(csv_data)

# changing the column names to all upper case
csv_data <- csv_data %>%
  rename_with(toupper)

# checking if all the bulbs changed to upper case
head(csv_data)
```

```{r, echo=TRUE}
# converting csv_data into tibble data, for easier data analysis
new_data <- as_tibble(csv_data)
new_data

# Extract rows by position
new_data %>%
  slice(1:40)

# Filter rows by some logical argument
csv_data %>%
  filter('PATIENT_PER_ICU_BED' > 3500)
```

```{r, echo=TRUE, include=TRUE}
# NEXT SET OF OUTPUT IS CREATED VIA INLINE CODE
```
This dataframe has `r nrow(csv_data)` rows and `r ncol(csv_data)` columns. The names of the columns and a brief description of each are in the table below:
`r names(new_data)`

```{r, echo=TRUE}
# Three important columns chosen are:
#     1. PATIENT_PER_ICU_BED
#     2. HIGH_RISK_PER_HOSPITAL
#     3. TOTAL_AT_RISK

total_summary <- summarise_each(csv_data[, c("PATIENT_PER_ICU_BED", "HIGH_RISK_PER_HOSPITAL", "TOTAL_AT_RISK")], funs(mean))

# publish summary
total_summary



```

